A \emph{Grimm} idea

### 2016-05-01 paper/main.tex
first paper version to present

\section{Pebble machines}
We are now ready to introduce the model of pebble machines formally. It closely
follows the notion of Turing machines but adds mechanisms to model the
movements of the agent. The presented model of pebble machines thus have
\begin{enumerate}
  \item a finite set of states $Q$
  \item a finite set of final states $F\subseteq Q$
  \item a finite set of distinguishable pebbles $P$
  \item a length of the used memory tape $m$
  \item a computation function $\delta$
  \item a function that describes the movement onto a vertex
    $\din$
  \item a function that describes the leaving of a vertex $\dout$
  \item an inital state $q_{0}$
\end{enumerate}
where formally the computation function models the behaviour of the pebble
machine as follows:
\begin{equation*}
  \delta\colon (Q\setminus F)\times\left\{0,1\right\}\rightarrow
  Q\times\left\{0,1\right\}\times\left\{L,R\right\}
\end{equation*}
with the typical understanding of the computation of a Turing machine.
As an agent traverses a vertex it is aware of the pebbles it carries
around, it observes the set of pebbles on the vertex and the degree of the
vertex and it furthermore aware of the port that it entered through.
Thus we use
\begin{equation*}
  \din\colon Q\times 2^{P}\times 2^{P}\times
  \left\{0,\dots,\Delta - 1\right\}\times\left\{0,\dots,\Delta - 1\right\}
  \rightarrow Q
\end{equation*}
to model this behaviour. And finally as the agent leaves a vertex it leaves
a set of pebbles behind, picks up another set and chooses the port it leaves
through. We choose to model this behaviour as
\begin{equation*}
  \dout\colon Q
  \rightarrow 2^{P}\times 2^{P}\times \left\{0,\dots,\Delta - 1\right\}
\end{equation*}
Of course we also demand consistency of the pebbles, i.e. that if the agent
carried $P_{1}\subseteq P$ and found $P_{2}\subseteq P$ it is necessary that
$P_{1}$ and $P_{2}$ are disjoint and furthermore that if the agent leaves
carrying $P'_{1}$ and leaving $P'_{2}$ behind those sets are disjoint and
$P_{1}\cup P_{2} = P'_{1}\cup P'_{2}$.

@@

We can now find the $(\mathcal{O}(1),0,\mathcal{O}(\log n))$ pebble machine we
are looking for.
\begin{thm}
  \label{thm:pebblewalk}
  There is a $(\mathcal{O}(1),0,\mathcal{O}(\log n))$ pebble machine which
  explores every given graph $G$ with bounded degree $\Delta$ either completly
  or at least loops on $n$ distinct vertices.
\end{thm}
\begin{wrapfigure}{r}{0.6\textwidth}
  \caption{Illustration how to transform any graph to a 3-regular graph.}
  \label{fig:3reg}
  \resizebox{0.6\textwidth}{!}{\input{tikz/regsim}}
\end{wrapfigure}
\begin{proof}
  The main idea is to use the adapted algorithm $A$ to find for an
  universal exploration sequence for a 3-regular graph from above. Therefore
  we take $G$ and transform it to a 3-regular graph on which $A$ finds the
  corresponding universal exploration sequence. The transformation replaces
  any node $v$ with degree $d$ by $3d$ nodes which then form a circle where
  every nodes has degree $3$. The edges of $v$ are moved to nodes in the
  circle. Specifically the edge with port $i$ is replaced by 3 edges which are
  attached to $(v,i), (v,i+d), (v,i+2d)$. (Note that the other
  vertex attached to the edge is replaced by the same construction and therfore
  offers enough ports for the newly introduced edges.) Figure \ref{fig:3reg}
  illustrates this construction.  This simulation yields at most
  $3\Delta n^{2}$ many nodes and can be emulated with $\mathcal{O}(1)$ many
  memory space since it is only needed to store the second component of the
  inserted nodes which is bounded by $3\Delta$. Furthermore $A$ can be run on
  this constructed Graph and only needs $\mathcal{0}(\log n)$ space since
  $\mathcal{O}(\log(3\Delta n^{2})) = \mathcal{O}(\log(3) + \log(\Delta) +
  2\log(n)) = \mathcal{O}(\log n)$. Which yields the desired
  $(\mathcal{O}(1),0,\mathcal{O}(\log n))$ pebble machine.
\end{proof}

In order to obtain the desired boundary of $\log\log n$ space and $\log\log n$
pebbles we show that a pebble machine can be simulated by storing its tape
content by pebbles that are placed on a closed walk.
\begin{thm}
  \label{thm:simulation}
  For every graph $G$ a $(s,p,2m)$ pebble machine $M$ can be simulated by
  a $(cs,p + c,m)$ pebble machine that simulates the walk of $M$ or explores
  $G$.
\end{thm}
\begin{figure}
  \caption{Illustration how to simulate the tape of a pebble machine by
  placing pebbles on a closed walk where $m_{1} = 3$. Note that the first
  $2^{m_{1}}=2^{3}=8$ distinct vertices are used for the simulation but the
  walk can be longer (as indicated by the dotted connection from $7$ to $0$).}
  \label{fig:tapesim}
  \begin{center}
    \resizebox{0.6\textwidth}{!}{\input{tikz/tapesim}}
  \end{center}
\end{figure}
\begin{proof}
  In this proof we will focus on the simulation and we leave out a thorough
  analysis of the space requirements of the simulation since it is rather
  technical but does not add a lot to the idea. Nevertheless we encourage
  interested readers to see \cite{pebbles} for the details.

  The main idea of this simulation is to store the tape of $M$ by pebbles
  placed on a closed walk in the graph. We show that every node on this closed
  walk can be given an index and a pebble placed on a node represents the
  binary representation of the index as the content of $M$'s tape. Therefore we
  first off introduce a variable $m_{1}$ which is dependent on the simulated
  pebble machine $M$ (further requirements on $m_{1}$ and how to obtain it
  can be seen in \cite{pebbles}). We will now take the tape of $M$ and separate
  it into $2m/m_{1}$ blocks of the size $m_{1}$. And each of these blocks is
  encoded by the placement of a pebble on a closed walk $\omega$. So we use
  pebble $p_{0}$ to encode the first $m_{1}$ bits, the pebble $p_{1}$ used to
  encode the next $m_{1}$ bits and so forth. Thus, the pebble $p_{i}$ encodes
  the bits $im_{1},\dots,(i+1)m_{1}-1$. It is therefore necessary that every
  pebble can encode $2^{m_1}$ different states. Therefore we use a closed walk
  $\omega$ which visits at least $2^{m_1}$ distinct vertices yielding $2^{m_1}$
  many \enquote{states} for each pebble (by its placement on this closed walk).
  Therefore we can use the binary representation of the indices of the nodes
  on which the pebbles are placed as the content of the encoded memory block.
  Figure \ref{fig:tapesim} illustrates this.

@@

  $M_{\text{Walk}}$ can be implemented as $(\mathcal{O}(1),0,\mathcal{m_1})$

@@

The construction of barriers is inductively and therefore requires that the
base case can be constructed directly. Fortunately Theorem 4 in \cite{0barrier}
yields a graph which can be adapted by connecting two identical copies to a
0-barrier (for details we refer again to \cite{pebbles}). With this in mind we
introduce a gadget construction for a given $r$-barrier $B$ as it can be seen
in figure \ref{fig:gadget}\todo{add figure}. Replacing edges of a graph with
such gadgets enforces, in a sense, \enquote{locality} of pebbles. We call
the verices that are connected by the gadget macro-vertices and if a $r+1$
pebble machine stands on a macro vertex we can safely be assured that all
pebbles are either carried by the agent or are placed within the adjacent
gadgets because otherwise the agent has left a pebble across of an $r$-barrier
and crossed the barrier with $r$ pebbles contradicting that it is in fact an
$r$-barrier. This observation can be used to construct an $r$-barrier given
a $(r-1)$-barrier is known.
\begin{thm}
  With a given $(r-1)$-barrier for an agent $A$ a $r$-barrier can be
  constructed.
\end{thm}
\begin{proof}
  We will present only the key ideas of this proof since it involves a
  few technicalities which do not add much to the general understanding
  and are therefore omitted. Given a subset of pebbles of cardinality $r$
  we can construct an $r$-barrier for this subset by using the given
  $(r-1)$-barrier to enforce locality of the pebbles. We can then project
  $A$'s behaviour to an agent $B$ which is a pebble machine without pebbles
  and has a state for every configuration (which includes the placing of the
  pebbles as well as state and tape content) of $A$. $B$ now moves over the
  macro-vertices as $A$ moves through the gadgets. But $B$ has no pebbles and
  thus, it is applicable to the 0-barrier. Projecting this construct back to
  the gadget construction yields an $r$-barrier for $A$ and the chosen subset
  of pebbles. Connecting now such $r$-barriers for all possible subsets of
  pebbles gives the desired $r$-barrier because $A$ always fails to traverse
  the $r$-barrier particular constructed for the pebble set used by $A$.
\end{proof}
By constructing such barriers it is possible to find for every agent
with $s$ states and $p$ pebbles (note that here the tape of the agent is
encoded in the states) it is possible to give a trap where the amount of
vertices can be bound by $\mathcal{O}(s^{8^{p+1}})$. Which can be used to
show that every agent with sub-logarithmic bits of memory needs at least
$\Omega(\log\log n)$ many pebbles to explore at least $n$ vertices.

>>>>

\section{Pebble machines}
We are now ready to introduce the model of pebble machines. It closely
follows the notion of Turing machines but adds mechanisms to model the
movements of the agent. The presented model of pebble machines thus have
\begin{enumerate}
  \item a finite set of states $Q$
  \item a finite set of final states $F\subseteq Q$
  \item a finite set of distinguishable pebbles $P$
  \item a length of the used memory tape $m$
  \item a computation function $\delta$
  \item a function that describes the movement onto a vertex
    $\din$
  \item a function that describes the leaving of a vertex $\dout$
  \item an inital state $q_{0}$
\end{enumerate}
where formally the computation function models the behaviour of the pebble
machine as follows:
\begin{equation*}
  \delta\colon (Q\setminus F)\times\left\{0,1\right\}\rightarrow
  Q\times\left\{0,1\right\}\times\left\{L,R\right\}
\end{equation*}
with the typical understanding of the computation of a Turing machine.
As an agent traverses a vertex it is aware of the pebbles it carries
around, it observes the set of pebbles on the vertex and the degree of the
vertex and it furthermore aware of the port that it entered through.
Thus we use
\begin{equation*}
  \din\colon Q\times 2^{P}\times 2^{P}\times
  \left\{0,\dots,\Delta - 1\right\}\times\left\{0,\dots,\Delta - 1\right\}
  \rightarrow Q
\end{equation*}
to model this behaviour. And finally as the agent leaves a vertex it leaves
a set of pebbles behind, picks up another set and chooses the port it leaves
through. We choose to model this behaviour as
\begin{equation*}
  \dout\colon Q
  \rightarrow 2^{P}\times 2^{P}\times \left\{0,\dots,\Delta - 1\right\}
\end{equation*}
Of course we also demand consistency of the pebbles, i.e. that if the agent
carried $P_{1}\subseteq P$ and found $P_{2}\subseteq P$ it is necessary that
$P_{1}$ and $P_{2}$ are disjoint and furthermore that if the agent leaves
carrying $P'_{1}$ and leaving $P'_{2}$ behind those sets are disjoint and
$P_{1}\cup P_{2} = P'_{1}\cup P'_{2}$.

@@

We can now find the $(\mathcal{O}(1),0,\mathcal{O}(\log n))$ pebble machine we
are looking for.
\begin{thm}
  \label{thm:pebblewalk}
  There is a $(\mathcal{O}(1),0,\mathcal{O}(\log n))$ pebble machine which
  explores every given graph $G$ with bounded degree $\Delta$ either completly
  or at least loops on $n$ distinct vertices.
\end{thm}
\begin{wrapfigure}{r}{0.5\textwidth}
  \caption{Illustration how to transform any graph to a 3-regular graph.}
  \label{fig:3reg}
  \resizebox{0.5\textwidth}{!}{\input{tikz/regsim}}
\end{wrapfigure}
\begin{proof}
  The main idea is to use the adapted algorithm $A$ to find for an
  universal exploration sequence for a 3-regular graph from above. Therefore
  we take $G$ and transform it to a 3-regular graph on which $A$ finds the
  corresponding universal exploration sequence. The transformation replaces
  any node $v$ with degree $d$ by $3d$ nodes which then form a circle where
  every nodes has degree $3$. The edges of $v$ are moved to nodes in the
  circle. Specifically the edge with port $i$ is replaced by 3 edges which are
  attached to $(v,i), (v,i+d), (v,i+2d)$. (Note that the other
  vertex attached to the edge is replaced by the same construction and therfore
  offers enough ports for the newly introduced edges.) Figure \ref{fig:3reg}
  illustrates this construction.  This simulation yields at most
  $3\Delta n^{2}$ many nodes and can be emulated with $\mathcal{O}(1)$ many
  memory space since it is only needed to store the second component of the
  inserted nodes which is bounded by $3\Delta$. Furthermore $A$ can be run on
  this constructed Graph and only needs $\mathcal{0}(\log n)$ space since
  $\mathcal{O}(\log(3\Delta n^{2})) = \mathcal{O}(\log(3) + \log(\Delta) +
  2\log(n)) = \mathcal{O}(\log n)$. Which yields the desired
  $(\mathcal{O}(1),0,\mathcal{O}(\log n))$ pebble machine.
\end{proof}

In order to obtain the desired boundary of $\log\log n$ space and $\log\log n$
pebbles we show that a pebble machine can be simulated by storing its tape
content by pebbles that are placed on a closed walk.
\begin{thm}
  \label{thm:simulation}
  For every graph $G$ a $(s,p,2m)$ pebble machine $M$ can be simulated by
  a $(cs,p + c,m)$ pebble machine that simulates the walk of $M$ or explores
  $G$.
\end{thm}
\begin{proof}
  In this proof we will focus on the simulation and we leave out a thorough
  analysis of the space requirements of the simulation since it is rather
  technical but does not add a lot to the idea. Nevertheless we encourage
  interested readers to see \cite{pebbles} for the details.

  The main idea of this simulation is to store the tape of $M$ by pebbles
  placed on a closed walk in the graph. We show that every node on this closed
  walk can be given an index and a pebble placed on a node represents the
  binary representation of the index as the content of $M$'s tape. Therefore we
  first off introduce a variable $m_{1}$ which is dependent on the simulated
  pebble machine $M$ (further requirements on $m_{1}$ and how to obtain it
  can be seen in \cite{pebbles}). We will now take the tape of $M$ and separate
  it into $2m/m_{1}$ blocks of the size $m_{1}$. And each of these blocks is
  encoded by the placement of a pebble on a closed walk $\omega$. So we use
  pebble $p_{0}$ to encode the first $m_{1}$ bits, the pebble $p_{1}$ used to
  encode the next $m_{1}$ bits and so forth. Thus, the pebble $p_{i}$ encodes
  the bits $im_{1},\dots,(i+1)m_{1}-1$. It is therefore necessary that every
  pebble can encode $2^{m_1}$ different states. Therefore we use a closed walk
  $\omega$ which visits at least $2^{m_1}$ distinct vertices yielding $2^{m_1}$
  many \enquote{states} for each pebble (by its placement on this closed walk).
  Therefore we can use the binary representation of the indices of the nodes
  on which the pebbles are placed as the content of the encoded memory block.
  Figure \ref{fig:tapesim} illustrates this.
  \begin{figure}[h!]
    \caption{Illustration how to simulate the tape of a pebble machine by
      placing pebbles on a closed walk where $m_{1} = 3$. Note that the first
      $2^{m_{1}}=2^{3}=8$ distinct vertices are used for the simulation but the
    walk can be longer (as indicated by the dotted connection from $7$ to $0$).}
    \label{fig:tapesim}
    \begin{center}
      \resizebox{0.6\textwidth}{!}{\input{tikz/tapesim}}
    \end{center}
  \end{figure}

@@

  $M_{\text{Walk}}$ can be implemented as $(\mathcal{O}(1),0,\mathcal{O}(m_{1}))$

@@

\begin{wrapfigure}{r}{0.4\textwidth}
  \caption{Gadget construction that ensures locality of pebbles.}
  \label{fig:gadget}
  \resizebox{0.4\textwidth}{!}{\input{tikz/gadget}}
\end{wrapfigure}
The construction of barriers is inductively and therefore requires that the
base case can be constructed directly. Fortunately Theorem 4 in \cite{0barrier}
yields a graph which can be adapted to a 0-barrier. With this in mind we
introduce a gadget construction for a given $r$-barrier $B$ as it can be seen
in Figure \ref{fig:gadget}. Replacing edges of a graph with
such gadgets enforces, in a sense, \enquote{locality} of pebbles. We call
the verices that are connected by the gadget macro-vertices and if a $r+1$
pebble machine stands on a macro vertex we can safely be assured that all
pebbles are either carried by the agent or are placed within the adjacent
\enquote{edges} because otherwise the agent has left a pebble across of an
$r$-barrier and crossed the barrier with $r$ pebbles contradicting that it is
in fact an $r$-barrier. This observation can be used to construct an
$r$-barrier given a $(r-1)$-barrier is known.
\begin{thm}
  With a given $(r-1)$-barrier for an agent $A$ a $r$-barrier can be
  constructed.
\end{thm}
\begin{proof}
  We will present only the key ideas of this proof since it involves a
  few technicalities which do not add much to the general understanding
  and are therefore omitted. Given a subset of pebbles of cardinality $r$
  we can construct an $r$-barrier for this subset by using the given
  $(r-1)$-barrier. By the above described gadget construction we can enforce
  the locality of the used $r$ pebbles. We can then project
  $A$'s behaviour to an agent $B$ which is a pebble machine without pebbles
  and has a state for every configuration (which includes the placing of the
  pebbles as well as state and tape content) of $A$. $B$ now moves over the
  macro-vertices as $A$ moves through the gadgets. But $B$ has no pebbles and
  thus, it is applicable to the 0-barrier. Projecting this construct back to
  the gadget construction yields an $r$-barrier for $A$ and the chosen subset
  of pebbles. Connecting now such $r$-barriers for all possible subsets of
  pebbles gives the desired $r$-barrier because $A$ always fails to traverse
  the $r$-barrier particularly constructed for the pebble set used by $A$.
\end{proof}
By constructing such barriers it is possible to find for every agent
with $s$ states and $p$ pebbles (note that here the tape of the agent is
encoded in the states) it is possible to give a trap where the amount of
vertices can be bound by $\mathcal{O}(s^{8^{p+1}})$. Which can be used to
show that every agent with sub-logarithmic bits of memory needs at least
$\Omega(\log\log n)$ many pebbles to explore at least $n$ vertices.

\section{Conclusion}
In this paper re-presented the idea of pebble machines for graph exploration
as discussed in \cite{pebbles}. For this we focused on conwaying the essential
ideas and constructions rather than explore the technical details. The main
result clearly is the simulation of memory of pebble machines by encoding
memory blocks by pebbles and that these pebbles keep a state by their placement
along a closed walk.
